---
title: "P8105: Data Science I"
author: "Assignment 3<br>Zhuodiao Kuang (UNI: zk2275)"
output:
  github_document:
    toc: TRUE
always_allow_html: TRUE
---

```{r load libraries,include=FALSE}
library(tidyverse)
library(patchwork)
library(ggridges)
library(ggplot2)
library(dplyr)
library(knitr)
library(gt)
library(kableExtra)
```



```{r setup, echo = FALSE, cache = FALSE, message = FALSE}
rm(list = ls())

## Global options
opts_chunk$set(echo      = TRUE,
	             cache     = TRUE,
               prompt    = FALSE,
               tidy      = FALSE,
               comment   = NA,
               message   = FALSE,
               warning   = FALSE,
               dpi       = 150,
               fig.width = 6, 
               fig.height = 12,
               fig.asp   = 5/7,
               fig.align = "center")


theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1


```{r include=FALSE}
library(p8105.datasets)
data("instacart")
```



```{r eval=FALSE}
total_aisles <- instacart|> pull(aisle)|>unique()|>length()
popular_aisles <- instacart |>
  group_by(aisle) |>
  summarise(item_count = n()) |>
  arrange(-item_count) # descending order
```


There are total of 134 aisles in this dataset, and "fresh vegetables" aisle is the most items ordered from.


```{r}
filtered_aisles <- instacart |>
  group_by(aisle) |>
  summarise(item_count = n()) |>
  filter(item_count > 10000) |> #limiting this to aisles with more than 10000 items ordered
  arrange(item_count)
filtered_aisles |> ggplot(aes(x = reorder(aisle, -item_count, FUN=mean), y = item_count)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Items Ordered from Each Aisle",
       x = "Aisle",
       y = "Number of Items Ordered")
```


This plot displays the number of items ordered in each aisle that with more than 10000 items ordered, among all, fresh fruits and fresh vegetables have the most orders. 

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  count(aisle, product_name, sort = T) |> 
  group_by(aisle) |> 
  top_n(3) |> # top 3 products within each group
  kable()
```

This table shows the the number of times each item is ordered in three most popular items from three aisles (“baking ingredients”, “dog food care”, and “packaged vegetables fruits”).


```{r, warning=FALSE, message=FALSE}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  group_by(product_name,order_dow) |> 
  summarise(mean_hour = mean(order_hour_of_day))|>  # 3 cols we focus on
  pivot_wider(
    names_from = product_name,
    values_from = mean_hour
  ) |> 
  mutate(order_dow = factor(order_dow,
                            labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))|> 
  kable(format = "markdown", 
        col.names = c("Day of Week", "Pink Lady Apples (Mean Hour)", "Coffee Ice Cream (Mean Hour)"),
        caption  = "Mean Order Hour for Selected Items") 

```



This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered throughout a week. 

### Problem 2


```{r message=FALSE, warning=FALSE, echo=TRUE}
library(p8105.datasets)
data("brfss_smart2010")
```

First, do some data cleaning:

```{r}
cleaned_brfss <- brfss_smart2010 |>
  janitor::clean_names() |>
  rename(State = locationabbr)|>
  filter(
    topic == "Overall Health", 
    response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")
    ) |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )
```


```{r message=FALSE, warning=FALSE}
states_observed <- cleaned_brfss |>
  group_by(year, State) |>
  summarise(Locations_Count = n_distinct(locationdesc), .groups = "drop") |>
  filter(Locations_Count >= 7)|>
  arrange(desc(Locations_Count))

# In 2002
states_2002 <- states_observed |> filter(year == 2002) |>select(State,Locations_Count)
kable(states_2002,booktabs = TRUE,
      caption  = "States that observed at 7 or more locations in 2002")|>
  kable_styling(position = "center",full_width = F)


# In 2010
states_2010 <- states_observed |> filter(year == 2010) |>select(State,Locations_Count)
kable(states_2010, booktabs=TRUE,
      caption  = "States that observed at 7 or more locations in 2010")|>
  kable_styling(position = "center",full_width = F)

```


In 2002, PA, MA, NJ, CT, FL, NC (total of 6 states) were observed at 7 or more locations. In 2010, total of 14 states were observed at 7 or more locations, including FL, NJ, TX, CA, MD, NC, NE, WA, MA, NY, OH, CO, PA, SC. 


```{r}
excellent_responses <- cleaned_brfss |>
  filter(response == "Excellent") |>
  group_by(year, State) |>
  summarise(average_value = mean(data_value, na.rm = TRUE), .groups = "drop")

ggplot(excellent_responses, aes(x = year, y = average_value, group = State, color = State)) +
  geom_line() +
  theme_minimal()+ 
  theme(legend.position = "right") +
  # add meta-data
  labs(
      title    = "Crude Prevalence over Time"
    , subtitle = "for each state"
    , x        = "Year"
    , y        = "Average Crude Prevalence Score"
    , caption  = "Source: BRFSS."
  )
```


This “spaghetti” plot contains only "Excellent" responses, years, states, and the average values across locations within a state.


```{r}
cleaned_brfss |>
  filter(State == "NY", year %in% c(2006, 2010))|>
# group by year
  group_by(year) |>
  # instantiate plot
  ggplot(aes(x = response, y = data_value, color = response)) +
  # add box plot
  geom_boxplot() +
  # stratify by year
  facet_grid(~year) + 
  # omit legend
  theme(legend.position = "none") +
  # add meta-data
  labs(
      title    = "Distribution of Crude Prevalence"
    , subtitle = "in NY State for given year and for given response type"
    , x        = "Response Type"
    , y        = "Crude Prevalence Score"
    , caption  = "Source: BRFSS."
  )
```


This two-panel plot shows the distribution of 'data_value' for four levels of responses in years 2006 and year 2010, among locations in NY State only. Year 2010 has an overall high data values than year 2006 in all responses. 

### Problem 3
```{r message=FALSE, warning=FALSE}
#Load, tidy, merge, organize datasets
covar <- read_csv("nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
      sex = recode(sex, "1" = "male", "2" = "female"),
      education = recode(education, "1" = "Less than high school", "2" = "High school equivalent", "3" = "More than high school")
      ) |>
  mutate(
    sex = factor(sex), 
    education = factor(education)
    )
clean_covar <- covar |>
  filter(age >= 21) |>
  drop_na()
accel <- read_csv("nhanes_accel.csv") |>
  janitor::clean_names() 
merged_data <- merge(clean_covar, accel, by = "seqn")
```


The "merged_data" dataset includes all variables from both 'accel' and 'covar' which merged by "SEQN" number; no data of participants less than age of 21; no missing demographic data; and all numeric values changed to meaningful factors. 


```{r}
# Table of the number of men and women in each education category
gender_edu <- merged_data |>
  group_by(sex, education) |>
  summarise(count = n(), .groups = "drop") |>
  pivot_wider(names_from = sex, values_from = count)
kable(gender_edu, 
      title = "Number of men and women in each education levels",
      col.names = c("Education Level", "Number of Female", "Number of Male"))
# Visualization of Age Distribution
ggplot(merged_data, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Age Distribution by Gender and Education",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


The table shows there are more female in the higher education group, more male have a high school equivalent background. This box-plot shows the age distribution of men and women in three different education levels. As a result, younger people appear to have a higher education level (more than high school), the middle age group have relevant lower education.


```{r message=FALSE, warning=FALSE}
new_merged <- merged_data |>
  mutate(
    total_activity = rowSums(across(min1:min1440), na.rm = TRUE)
    )
ggplot(new_merged, aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = .5) +
  facet_wrap(~education) +
  geom_smooth() +
  labs(title = "Total Activity by Age, Gender, and Education")
```


The y-axis in this panel plot represents the overall activities, and the x-axis represents the ages. In order to compare men and women and their activities, it divides into three subplots based on educational attainment. Consequently, there is a direct correlation between activities and age; individuals who are older (over 60) tend to be less active. There seems to be more activity and a gentler line in the group with higher education levels.  


```{r}
activity_data <- merged_data |>
  group_by(education, sex)|>
  summarise(across(starts_with("min"), ~ mean(.), .names = "mean_{.col}"), .groups = "drop") |>
  pivot_longer(cols = starts_with("mean_"), names_to = "time", values_to = "mean")|>
  mutate(
    time = substring(time, 9),
    time = as.numeric(time)
  )
ggplot(activity_data, aes(x = time, y = mean, color = sex)) +
  geom_line() +
  facet_grid(. ~education) +
  labs(title = "24-Hour Activity by Education and Gender", 
       x = "Time in mins",
       y = "Mean value of activity")
```


The 24-hour activity time courses for three educational levels are displayed in this three-panel figure, along with a gender correlation analysis. Because of this, the "less than high school" group has a notably greater mean value of activity than the other education level groups, and female tends to have higher activity score than male except for the “High school equivalent” education level.