---
title: "p8105_hw3_zk2275"
author: "Zhuodiao Kuang"
date: "`r Sys.Date()`"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(patchwork)
library(ggridges)
library(ggplot2)
library(dplyr)
library(knitr)
library(gt)
```


```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1


```{r include=FALSE}
library(p8105.datasets)
data("instacart")
```



```{r eval=FALSE}
total_aisles <- length(unique(instacart$aisle))
popular_aisles <- instacart |>
  group_by(aisle) |>
  summarise(item_count = n()) |>
  arrange(-item_count)
```


There are total of 134 aisles in this dataset, and "fresh vegetables" aisle is the most items ordered from.


```{r}
filtered_aisles <- instacart |>
  group_by(aisle) |>
  summarise(item_count = n()) |>
  filter(item_count > 10000) |>
  arrange(item_count)
ggplot(data = filtered_aisles, aes(x = aisle, y = item_count)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Items Ordered from Each Aisle",
       x = "Aisle",
       y = "Number of Items Ordered")
```


This plot displays the number of items ordered in each aisle that with more than 10000 items ordered, among all, fresh fruits and fresh vegetables have the most orders. 


```{r, message=FALSE, warning=FALSE}
selected_aisles <- instacart |>
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") |>
  group_by(aisle, product_name) |>
  summarise(order_count = n(), .groups = "drop") |>
  arrange(aisle, -order_count) |>
  filter(min_rank(desc(order_count)) < 4)
kable(selected_aisles, 
      title = "Top 3 Items in Selected Aisles", 
      col.names = c("Aisle", "Product Name", "Order Count"))
```


This table shows the the number of times each item is ordered in three most popular items from three aisles (“baking ingredients”, “dog food care”, and “packaged vegetables fruits”).


```{r, warning=FALSE, message=FALSE}
selected_items <- instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarise(mean_hour = mean(order_hour_of_day), .groups = "drop") |>
  pivot_wider(names_from = product_name, values_from = mean_hour) |>
  mutate(order_dow = factor(order_dow,
                            labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))
selected_items |> 
  kable(format = "markdown", 
        col.names = c("Day of Week", "Pink Lady Apples (Mean Hour)", "Coffee Ice Cream (Mean Hour)"),
        title = "Mean Order Hour for Selected Items") 
```



This table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered throughout a week. 

